---
title: "Batería de pruevas"
format: html
editor: 
  markdown: 
    wrap: 72
---

# Required Libraries

library(tidyverse) library(cluster)

# Define your dataset and parameters

data \<- your_dataset number_of_clusters \<- k \# Define your 'k' value

# Initial Clustering (e.g., using K-means)

initial_clusters \<- kmeans(data, centers = number_of_clusters)

# Function to calculate silhouette scores

calculate_silhouette \<- function(data, clusters) { \# Compute
silhouette scores \# Return the scores }

# Recursive Elimination Function

recursive_elimination \<- function(data, threshold) { \# Implement the
elimination logic \# Return the reduced dataset and average silhouette
scores }

# Plotting the results

plot_silhouette_vs_samples \<- function(scores) { \# Use ggplot2 to
create the plot }

# Execution of the process

scores \<- recursive_elimination(data, 2 \* number_of_clusters)
plot_silhouette_vs_samples(scores)

# Document the process in RMarkdown/Quarto

```{r, message=FALSE, warning=FALSE}
# Required Libraries
set.seed(42)
source("../INPUT/FUNCTIONS/preprocess.R")
library(dplyr)
library(rsample)
library(recipes)
library(randomForest)
library(corrplot)
library(tibble)
library(purrr)
library(reshape2)
library(ggplot2)
library(plotly)
library(scatterplot3d)
library(factoextra)
library(cluster)
library(tidyverse)
```

```{r}
data <- read.csv('../Input/DATA/Summary_Dataset_Multivar_Percentage_MAU.csv')
data_multi <- read.csv('../Input/DATA/Summary_Dataset_Multivar_Relative_MNAU.csv')
```

```{r}
data = preprocess(data)
row.names(data)
```

```{r}
#split <- split_train_SH(data)

#data <- split$train_data
#sh <- split$sh_data


# Store original row names of the datasets
data_row_names <- rownames(data)
#sh_row_names <- rownames(sh)
data_col_names <- colnames(data)
#sh_col_names <- colnames(sh)

# Step 1: Create a recipe
recipe <- recipe(~ ., data = data) %>%
  step_normalize(all_predictors(), -all_of(c("Type", "AccumulationType", "Individuals")))

# Step 2: Fit the recipe to the training set
fitted_recipe <- prep(recipe, training = data)

# Step 3: Apply the transformation to both training and test sets
data_transformed <- bake(fitted_recipe, new_data = data)
#sh_transformed <- bake(fitted_recipe, new_data = sh)

# Reattach the row names
rownames(data_transformed) <- data_row_names
#rownames(sh_transformed) <- sh_row_names
col.names(data_transformed) <- data_col_names
#col.names(sh_transformed) <- sh_col_names

data <- data_transformed
sh <- sh_transformed
```

```{r}
library(cluster)
library(broom)
set.seed(42)
cols_km <- setdiff(names(data), c("Type","AccumulationType","Individuals", "Cluster_Pnas"))
data_km <- data[cols_km]
```

```{r}
library(cluster)
library(broom)
set.seed(42)
cols_km <- setdiff(names(data), c("Type","AccumulationType","Individuals", "Cluster_Pnas"))
data_km <- data[cols_km]

# Load necessary libraries
library(factoextra)
library(cluster)

#Perform K-means clustering
set.seed(42)  
kmeans_result <- kmeans(data_km, centers = 8, nstart = 25)  

# The cluster assignments are stored in kmeans_result$cluster
sub_grp <- kmeans_result$cluster

# Visualize the clusters
fviz_cluster(list(data = data_km, cluster = sub_grp))
```

```{r}
calculate_silhouette <- function(data, clusters) {
  # Convert cluster assignments to numeric
  cluster_vector <- as.numeric(as.character(clusters$cluster))

  # Ensure the data passed to dist() is numeric
  numeric_data <- data %>% select_if(~is.numeric(.))

  # Compute silhouette scores
  sil_scores <- silhouette(cluster_vector, dist(numeric_data))

  # Extract the silhouette width values
  sil_widths <- sil_scores[, 'sil_width']

  return(sil_widths)
}



```

```{r}
calculate_silhouette(data_km,clusters=clusters)
```

```{r}
 # Initial Clustering (for example, using K-means)
  k <- 3  # Example number of clusters, adjust as needed
  clusters <- kmeans(data_km, centers = k)

  # Variable to hold average silhouette scores
  avg_sil_scores <- numeric()
```

```{r}
recursive_elimination <- function(data, threshold, number_of_clusters) {
  require(dplyr)

  # Initial Clustering (for example, using K-means)
  k <- number_of_clusters  # Example number of clusters, adjust as needed
  clusters <- kmeans(data, centers = k)

  # Variable to hold average silhouette scores
  avg_sil_scores <- numeric()

  while (nrow(data) > threshold) {
    # Calculate silhouette scores
    sil_widths <- calculate_silhouette(data, clusters)
    
    # Record the average silhouette score
    avg_sil_scores <- c(avg_sil_scores, mean(sil_widths))
    
    # Identify the sample with the lowest silhouette score
    worst_sample_idx <- which.min(sil_widths)
    
    # Remove the worst sample
    data <- data[-worst_sample_idx, ]
    
    # Update clustering with the reduced dataset
    clusters <- kmeans(data, centers = k)
  }

  return(list(reduced_data = data, avg_silhouette_scores = avg_sil_scores))
}

```

```{r}
library(ggplot2)

plot_silhouette_vs_samples <- function(scores) {
  # Creating a data frame for ggplot
  data_to_plot <- data.frame(
    SampleSize = length(scores):1,  # Assuming each step reduces the sample size by 1
    AvgSilhouetteScore = scores
  )

  # Plotting using ggplot
  ggplot(data_to_plot, aes(x = SampleSize, y = AvgSilhouetteScore)) +
    geom_line() +  # Line plot
    geom_point() +  # Points to show each step
    labs(title = "Average Silhouette Score vs. Sample Size",
         x = "Sample Size",
         y = "Average Silhouette Score") +
    theme_minimal()
}

```

# Recursive sample elimination maximizing the Silhouette Score

```{r}
# Load required libraries
library(cluster)
library(ggplot2)
library(dplyr)

# Define your dataset and parameters
data <- data_for_clustering # Replace with your actual data
number_of_clusters <-  # Define your 'k' value

# Function to calculate silhouette scores
calculate_silhouette <- function(data, clusters) {
  cluster_vector <- as.factor(clusters$cluster)
  sil_scores <- silhouette(cluster_vector, dist(data))
  sil_widths <- sil_scores[, 'sil_width']
  return(sil_widths)
}

# Recursive Elimination Function
recursive_elimination <- function(data, threshold,  number_of_clusters) {
  clusters <- kmeans(data, centers = number_of_clusters)
  avg_sil_scores <- numeric()

  while (nrow(data) > threshold) {
    sil_widths <- calculate_silhouette(data, clusters)
    avg_sil_scores <- c(avg_sil_scores, mean(sil_widths))
    worst_sample_idx <- which.min(sil_widths)
    data <- data[-worst_sample_idx, ]
    clusters <- kmeans(data, centers = number_of_clusters)
  }

  return(list(reduced_data = data, avg_silhouette_scores = avg_sil_scores))
}

# Plotting the results
plot_silhouette_vs_samples <- function(scores) {
  data_to_plot <- data.frame(
    SampleSize = length(scores):1,
    AvgSilhouetteScore = scores
  )
  ggplot(data_to_plot, aes(x = SampleSize, y = AvgSilhouetteScore)) +
    geom_line() + geom_point() +
    labs(title = "Average Silhouette Score vs. Sample Size",
         x = "Sample Size", y = "Average Silhouette Score") +
    theme_minimal()
}

# Execution of the process
result <- recursive_elimination(data, 2 * number_of_clusters,  number_of_clusters)
scores <- result$avg_silhouette_scores
plot_silhouette_vs_samples(scores)
```

```{r}
# Load required libraries
library(cluster)
library(ggplot2)
library(dplyr)

# Define your dataset and parameters
data <- data_for_clustering # Replace with your actual data
initial_sample_size <- nrow(data)
number_of_clusters <-  3# Define your 'k' value

calculate_silhouette <- function(data, clusters) {
  # Convert cluster assignments to numeric
  cluster_vector <- as.numeric(as.character(clusters$cluster))

  # Ensure the data passed to dist() is numeric
  numeric_data <- data %>% select_if(~is.numeric(.))

  # Compute silhouette scores
  sil_scores <- silhouette(cluster_vector, dist(numeric_data))

  # Extract the silhouette width values
  sil_widths <- sil_scores[, 'sil_width']

  return(sil_widths)
}

# Recursive Elimination Function
recursive_elimination <- function(data, threshold) {
  clusters <- kmeans(data, centers = number_of_clusters)
  avg_sil_scores <- numeric()

  while (nrow(data) > threshold) {
    sil_widths <- calculate_silhouette(data, clusters)
    avg_sil_scores <- c(avg_sil_scores, mean(sil_widths))
    worst_sample_idx <- which.min(sil_widths)
    data <- data[-worst_sample_idx, ]
    clusters <- kmeans(data, centers = number_of_clusters)
  }

  return(list(reduced_data = data, avg_silhouette_scores = avg_sil_scores))
}


plot_silhouette_vs_samples <- function(scores, initial_sample_size, k) {
  # Calculate the sequence for Sample Sizes
  # It starts from the initial sample size and ends at 2*k, decrementing by 1 each time
  sample_sizes <- seq(initial_sample_size, 2 * k, length.out = length(scores))

  # Creating the data frame for plotting
  data_to_plot <- data.frame(
    SampleSize = sample_sizes,
    AvgSilhouetteScore = scores
  )

  # Plotting using ggplot
  ggplot(data_to_plot, aes(x = SampleSize, y = AvgSilhouetteScore)) +
    geom_line() + geom_point() +
    labs(title = "Average Silhouette Score vs. Sample Size",
         x = "Sample Size", y = "Average Silhouette Score") +
    theme_minimal()
}


# Execution of the process
result <- recursive_elimination(data_for_clustering, 2 * number_of_clusters ,  number_of_clusters)
scores <- result$avg_silhouette_scores
plot_silhouette_vs_samples(scores, initial_sample_size, number_of_clusters)

```

```{r}
# Ensure data is suitable for clustering
# If data contains factors, either convert them to numeric or exclude them
data_for_clustering <- data_km %>%
  select_if(~is.numeric(.))  # This selects only numeric columns

# Proceed with clustering and analysis using data_for_clustering
initial_clusters <- kmeans(data_for_clustering, centers = number_of_clusters)

# Modify the rest of the code to use data_for_clustering instead of data
# ...

```

```{r}
scores
```
